---
title: "Should I Airbnb My Property?"
author: "Andy Krause and Gideon Aschwanden"
header-includes:
    - \usepackage{setspace}\doublespacing
date: "February 17, 2017"
output:
  pdf_document: default
  html_document: default
geometry: margin=3.3cm
bibliography: abb.bib
---

```{r load_libraries, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, results='hide'}

  library(spdep)
  library(maptools)
  library(gstat)
  library(ggplot2)
  library(geosphere)
  library(ggmap)
  library(xtable)
  library(chron)
  library(plyr)
  library(sp)
  library(rgeos)
  library(reshape2)
  library(stringr)
  library(RColorBrewer)
  library(Hmisc)
  library(kernlab)
  library(lmtest)
  library(ROCR)


```

```{r set_paths, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, results='hide'}

  # Get computer names

  comp.name <- Sys.info()['nodename']

  # Assign path based on computer name
  
  if(comp.name == '7020D-121777-W' | comp.name == 'DESKTOP-1D7JO4J'){
  
    data.path <- 'c:/dropbox/research/airBNB/data/'
    code.path <- 'c:/code/'
    fig.path <- 'c:/dropbox/research/airBNB/figures/'
    
  } else {
  
    data.path <- 'alternate data path'
    code.path <- 'alternate data path'
    
  }

 ## Set options

  options(xtable.comment = FALSE)
  options(xtable.caption.placement='top')
  
```

```{r source_files, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, results='hide'}

  source(paste0(code.path, "dataviztools/ggplottools.R"))
  source(paste0(code.path, "datamgmttools/dataMungetools.R"))
  source(paste0(code.path, "research/AirBNBMelbourne/abb_Functions.R"))

```

```{r set_colors, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE}

 # Set all colors

  abb.colors.names=c('rausch', 'hackberry', 'kazan', 'babu', 'lima', 'beach',
                     'ebisu', 'tirol', 'foggy', 'hof')
  abb.col <- c(rgb(250,88,99, maxColorValue=255), rgb(139,14,82, maxColorValue=255),
               rgb(0,117,140, maxColorValue=255), rgb(4,211,191, maxColorValue=255), 
               rgb(77,226,110, maxColorValue=255), rgb(252,179,14, maxColorValue=255),
               rgb(249,177,139, maxColorValue=255), rgb(191,166,100, maxColorValue=255),
               rgb(156,161,155, maxColorValue=255), rgb(86,94,97, maxColorValue=255))

  # Set specific colors
  
  str.col <- abb.col[1]
  ltr.col <- abb.col[3]
  
```

```{r load_workspace, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, cache=TRUE}

   load(paste0(data.path, '/analyzed/analyzed_workspace.RData'))

```

&nbsp;
&nbsp;

## Introduction

With the rise of Airbnb and other self-hosting accommodation platforms^[Flipkey.com, VRBO.com, Homeaway.com, Tripping.com as examples.], it is easier than ever for property owners and investors to consider short-term rental (STR) as a potential use for a revenue generating residential property. Should they though? Or rather, is it more profitable to do so when compared to a more traditional long term rental (LTR)? Are certain types of properties and certain locations, if any, more likely to show a preference for (in terms of profitability) short term leasing strategy over long-term.

To address these questions and to provide some guidance to potential and current investors and owners, we examine the residential rental market, both short- and long-term in Melbourne, Australia. Melbourne has a thriving short-term rental market with over 20,000 total properties listed on the Airbnb portal, the source of our short-term market data. We adopt a scenario-based approach that compares potential annual revenues for properties under both a short-term rental (STR) and a long-term rental (LTR) situation. We break the results down by property types (apartments vs houses), product type (e.g. 1 bed/1 bath) and by location (geographic submarket).  Our results shows that the feasibility of a greater return through a short term rental -- what we have termed ``Short Term Preference'' -- is dependent on property type, product and location.  In short, the answer to whether or not a owner or investor should opt for a short-term leasing strategy: it depends.  

## Literature Review

Existing research into short-term rental markets is sparse. However, with the arrival of Airbnb, and the controversies surrounding it, the topic is creating interest both in real estate and economics as well as in a wide variety of fields as evidenced by a growth in recent publications and on-going working papers.  Within the extant work, the focus remains centered on rural and/or amenity-based STRs such as beach-front and ski-resorts.  This focus resulted from the fact that the market itself operated almost wholly within this sphere.  The shift to small, urban dwellings within the STR market is relatively new and explains the dearth of past work in this direction. 

Much of the existing work focuses on price or rent formation in second home markets.  Specifically, the impacts of amenities such as water access [@nelson2010valuing], water quality [@gibbs2002hedonic; @clapper2014water], proximity to ski facilities  [@soguel2008impact; @nelson2010valuing] and theme parks [@tsai2015impact] are shown to positively impact sales prices in tourism-dominated areas.  Host policies can also matter as @benjamin2001value find that smoking prohibitions influence the weekly rental rates for vacation homes.

Focusing specifically on the short-term rental market, @cassidy2007tourist suggest that within Australia's Gold Coast, the STR industry lacks sophistication and efficiency.  Most notably, they find that most nightly or weekly rates are set by resident unit managers (RUMs) and are often under-priced as there is little accountability toward owners.  The authors also note that very little research has been put toward understanding this growing market segment.

A set of recent work has looked into the impact of short-term markets on housing prices and affordability. These working papers examine the influence of Airbnb on existing housing prices in New York [@sheppard2016airbnb], Los Angeles [@lee2016airbnb] and Boston [@merante2016home]. Though very broad analyses, all suggest that Airbnb has had a positive impact on local home prices and/or a negative impact on affordability. Other research examines the impact of short-term rentals on the hotel market, with @gutierrez2016airbnb finding that Airbnb and hotels have similar location patterns in Barcelona and @zervas2014rise showing an inverse relationship between Airbnb listings and hotel revenues.

A critical theme that arises out of the extant literature is that fundamentals of the markets for primary dwellings (owner-occupied and long-term rentals) may differ from those of vacation homes (or, by association, short-term rentals). Or, in other words, the factors that influence prices and rents in the two are not equivalent. @cho2003impacts find that the second home market is more sensitive to proximity to environmental amenities than the owner occupied market. In Switzerland, @soguel2008impact test to see if second homes pay a higher premium for ski resort access.  Their results fail to show a statistically significant difference in the impact of ski resorts between markets, but do find that other property characteristics have heterogenous effects between the two. In examining the difference between the urban and rural STR markets,  @nilsson2015influence makes a key distinction, finding that rural vacation home markets value amenities whereas urban holiday properties tend to derive large premiums from proximity to attractive residential areas.

Though issues of second homes were primarily treated as a rural phenomenom for decades  [@gallent2007second], a wave of ``new urban tourism'' @fuller2014stop has re-focused some of the problems of short term rentals on large, urban environments. With this move to urban areas, issues such as nuisance, property rights and regulation have naturally arisen.  As a result, the short-term rental market has garnered significant interest in the public policy and legal sphere as of late. As many second homes are not permitted for or regulated as commercial entities, STRs can have serious effects on neighbors in the way of externalies such as noise and congestion [@frost2006commentary].  While strict regulation is often proposed, doing so may constititute regulatory takings [@jefferson2015can].  Other, rights-based, regulatory approaches have been suggested [@miller2014transferable].  Many of the key legal decisions regarding short term rentals, particulary though Airbnb, are still in progress and much remains to be settled in this area in terms of the regulatory and legal environment in which STRs are allowed to operate.

Overall, the extant literature on short-term rentals is limited, but growing.  Of the three types of research -- pricing, impacts on housing market and legalities/externalities -- none specifically address the decision faced by an owner who is trying to decide between a long- or short-term leasing strategy. The closest analog in the literature is a piece by @larson2009pur examining the choice between purchasing a time-share or simply paying for hotel rooms.  Theoretically, the literature shows that in most cases there is a demonstratable difference between first and second home market dynamics due to the use differences of the occupants.  We hypothesize that this same thinking can be extended to the difference between short-term rental occupants (tourists and business travelers) and long-term rental occupants (workers and retirees) and, as a result, there should be differences in the market which make certain housing products and locations preferable (more profitable) under a short-term versus long-term tenure.

## Method

We approach answering the question of whether to pursue a short- or long-term leasing strategy by creating and assessing a scenario likely to be faced by a recent purchaser of a property.  More specifically, we examine the following case:

* An investor purchases a property in Melbourne, Australia on September 1, 2015.
* They are looking to maximize the net revenue over the 12-month period from September 1, 2015 to August 31, 2016.
* The choices are 1) traditional year long rental (long term), or: 2) short-term, nightly Airbnb-type rental.
* Only direct monetary revenues and costs are considered, others such as liquidity and inconvenience are not considered.
* As most long-term rental use a leasing agency, we will assume use of an agency. 


We will not, a priori, specify a location or property type for the scenario as we hypothesize that the preference of a short-term rental over a long-term one will vary by the location and property characteristics. The extent to which location, property type and property characteristics impact the probabilit of short-term preference are examined in the results section. 

Australia implements a particular tax policy, often termed `negative gearing', that allows for favourable tax treatment to real estate investors who experience greater costs than revenues on investment properties^[See @fane2004negative for a complete explanation of negative gearing.] To date, this policy only applies to long-term rental properties.  While this policy can influence the net revenue (post tax) of a property, the negative gearing policy requires knowledge of the investor's personal income situation.  As we do not have access to this information in the marketplace, we will not be considering any effects of this policy in our analysis here.

One benefit of a short-term rental may be the ability to adjust the nightly rate over the course of the year to reflect changing demand and market trends. As future market trends are unknown at the time of purchase and the decision of long- vs short-term leasing strategy is asymetrically irreversable (can go from short to long, but not long to short), we make our comparisons using a fixed nightly (short-term) rate to respresent the most likely state of knowledge for our hypothetical investor on September 1, 2015. Both long-term rents and short-term rates and occupancy structures will be based on the market conditions as of this date.

We do not directly observe properties that are utilized in both a short- and long-term leasing strategy. Therefore, comparisons are done based on estimated or imputed revenues. Most specifically, we will estimate the long-term revenue for observations of properties currently listed on a short-term rental portal (Airbnb). Initially, we attempted to also estimate short-term rental revenue for long-term properties but found the process of imputing occupancy rates and nightly rate structures too inaccurate to provide valid comparisons.

### Comparison

In this scenario we compare the observed 12-month revenues from short-term rental properties with their hypothetical long-term rental revenues.  We define revenue here as the gross revenue minus the direct costs associated with each of the two tenure possibilities. Costs which are equivalent between the two strategies are ignored.  

\begin{equation}
Rev_{ST} = RevG_{ST} - C_{ST}
\end{equation}

The gross short-term revenue is expressed by:

\begin{equation}
RevG_{ST} = Rate_{N} * O_{Total}
\end{equation}

where $Rate_{N}$ is the average nightly rate and $O_{Total}$ are the total number of occupancies (bookings) over the 12-month period. For long-term revenues:

\begin{equation}
Rev_{LT} = RevG_{LT} - C_{LT}
\end{equation}

Where $Rev_{LT}$ is the net revenue for a long-term rental, $RevG_{LT}$ is the gross revenue and $C_{LT}$ are the specific costs associated with long-term rental. The gross long-term revenue is expressed by:

\begin{equation}
RevG_{LT} = Rent_{M} * (12 - C_{S})
\end{equation}

where $Rent_{M}$ is the nominal monthly rent and $C_{S}$ are search costs defined in terms of the time (in months) spent to find a long-term tenant.

We have limited our consideration of costs in this case to be those undergone during a single year of ownership.  Broadly speaking, these fall into four categories long-term rental and five for short-term properties.  Long-term costs are summarized as:

\begin{equation}
C_{LT} = T + M + F + L
\end{equation}

where $T$ are tax related expenses, $M$ are general maintenance expenses, $F$ are financing costs and $L$ are leasing agency fees. Short-term properies experience three of the same costs ($T$, $M$ and $F$) plus additional janitorial costs $J$ between each of the short-term occupants as well as property utility costs, $U$ over the course of the year.

\begin{equation}
C_{ST} = T + M + F + J + U
\end{equation}

As our scenario compares a hypothetical and particular owner over the course of a single  year the tax $T$ and $F$ costs are identical between the two leasing strategy and therefore cancel out. General maintenance $M$ expenses may differ between the two strategies, and an argument could be made for either to be higher. Long-term renters spend more time in the property over the course of the year and bring their own appliances^[White goods are usually not supplied in the Australian rental market.] and furniture which may damage the home. They are also more likely to affix items to the structure.  Conversely, short-term rentals are usually furnished and the occupants have less incentive not to be evicted as their damage is not often uncovered until after they have left^[Note that Airbnb's rating policy may help protect hosts as bad guests can be flagged on the platform.].  Both arguments are meritous and we were unable to find any evidence to support maintenance costs being higher in short or long term rentals. As a result we have made the assumption that the maintenance costs will be identical, and they too cancel out.

Leasing costs $L$ are the fees paid to the rental agency to find and manage tenants over the life of the lease.  In Melbourne, rates generally range from 5\% to 7\% per annum.  We have assumed leasing costs of 6\% of the total gross revenue of the long-term lease contract. 

Janitorial costs, $J$, cover the costs to 'turn-over' the property between occupants.  For a short-term leasing strategy, these costs can be experienced every day or two, depending on the minimum length of stay and the total number of occupancies over the course of the year. Our data for short-term rentals comes from the Airbnb platform.  In addition to the nightly rate that an occupant pays to the owner there is also a separate 'cleaning fee' that is attached, as is a booking fee to Airbnb^[This fee could be considered analogous to the leasing fees $L$ that long-term rentals pay]. Some hosts add additional 'linen' fees to cover costs of laundering.  While it may be advantageous for hosts in competitive markets to underquote their cleaning fees to keep occupancies up, we have no evidence to suggest that owners are exhibiting such behavior. As a result of janitorial fees being a separate line item and directly paid by the occupants they can also be ignored in the comparative analysis below as they are direct.

Finally, there are utility costs, $U$, that are borne by the property owner in a short-term leasing strategy but by the tenant in a long-term option.  The State of Victoria estimates energy costs for an apartment at A\$330 per month and for a larger house at A\$520 per month.  For two people in an apartment the cost is approximately A\$5 per person per day and for a four person household in a detached house the cost is around A\$4 per person per day person. For the purposes of estimating utility costs for the short-term rentals we use a mid-point value of A\$4.50 per person per day.  As we don't know total number of guests in each reservation we use a figure of 1.7 guests per bedroom and sum up the total costs for all bookings over the year.   

\begin{equation}
U = (4.5 * B * 1.7) * O
\end{equation}

We will consider short-term rental (STR) strategy to be prefable to long-term rental (LTR) strategy if the following inequality holds:

\begin{equation}
Rev_{ST} > Rev_{LT}
\end{equation}

More specifically, if:

\begin{equation}
(Rate_{N} * O_{Total}) - U > (Rent_{M} * (12 - C_{S})) - L
\end{equation}

## Data

Data for this study were gathered from two different sources, one for short-term rental market and one for the long-term rental market.  Short-term rental market observations are from the Airbnb online portal. Airbnb is the largest short-term rental hub in the Melbourne region. As opposed to similar sites such as VRBO.com and Homeaway.com which focus on vacation homes in tourist dominated areas such as beaches and ski resorts, Airbnb offers considerable coverage in urban areas and, therefore, directly competes to a greater degree with more traditional long-term rentals than the other sites. The data on the Airbnb market in Melbourne have been purchased from www.airdna.co, a data provider specializing in Airbnb data collection and analysis. The short-term data include two types of observations, property level and nightly. Property-level data consists of a single observation for each property that is or has been listed on the Airbnb portal over the October 2014 to August 2016 period^[This time period represents the period in which AirDNA was engaged in data collection in Melbourne.]  The nightly data offer one observation for each property for each night of the year during its time on the portal.  These nightly data indicate the advertised price, the occupancy status and a unique reservation identification.

The Airbnb data cover the Melbourne metropolitan region from October 1st, 2014 to August 30th, 2016. There are missing nightly observations during the time period before September 1, 2015. After initially imputed these observations we found the results unreliable and have structured the analysis to only cover the time period containing complete data.  All monetary values in the Airbnb data are in US dollars and have been converted to Australian dollars using the median of the exchange rate during this time period, 1.32 Australian dollars to 1 US dollar.

The data on long-term leases was provided by Australia Property Monitors (APM), a Fairfax Group company. In raw form, this dataset includes an observation for each rental property listing for each time a detail of the listing changed. For example, if a property were listed for rent for one month and the price was dropped once each week there would be four observations in the dataset. These four observations are grouped by an event-specific identification number. This number is specific to the listing event, not the property. In each case the last observation signifies the final price of the rental contract. We separate this data into two sets for efficiency reasons. The first data set includes all information about the property, physical and economic (date and amount of rental contract), while the second contains a property ID, a rental event ID, the list price and the date of the change. As such, the first dataset is a set of market observations with associated characteristics and the second is a listing history of each rental event. The long-term lease data cover the period from January 2014 to December 2015.

Unfortunately, there are no common identifiers between the two datasets as the Airbnb data does not include address. Both sets of data do include latitude and longitude, but in a dense urban area with many multiple-family dwellings it is impossible to match apartments based on a two-dimensional coordinates. As a result of this, we are unable to directly match observations from the two datasets. However, the number of properties that rented long-term for the 2014 to 2015 period and then switched to Airbnb for the 2015 to 2016 period are likely very small and this inability to directly match properties does not greatly impact our modeling strategy.

```{r set_constants, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, results='hide'}

  exch.rate <- 1.32

```

### Preparation

Here we describe the operations that we have performed to prepare the two sets of raw data for analysis. Primarily, this involves the filtering out data errors, outlier values and other observations than contain data that do not fit our research questions. Additionally, as our data are from two different sources, a number of fields require standardization so that comparisons between the two leasing strategy options -- long-term rental (LTR) vs short-term rental (STR) -- can be made.

We begin by filtering observations by time. Our hypothetical leasing strategy decision is that of an investor who purchases a property on September 1, 2015 and must decide on traditional long-term rental versus an AirBNB-type, short-term rental approach to generating income. To determine the likely LTR rate for any given property, we will use the observed rental transactions from September 1, 2014 to August 31, 2015. Taking the Airbnb or STR approach allows the owner to change rates over time and is subjected to daily changes in the market (supply and demand factors). In this simulated example we will use AirBnB data from September 1, 2015 to August 31, 2016 to represent the actual AirBNB market conditions over the time period in question.

Next, we filter the data based on property types. AirBNB units are classified into 19 different property types. The most common are Apartments and Houses, the least common are Igloo, Tent, Treehouse and Yurt^[Property types are owner/lister defined, therfore it is likely that some designation, such as Igloo, at the very least, are not completely true representations of the property type]. For the sake of this analysis, we will collapse these twenty types into three types:

1. **House**:  Includes properties labeled 'House' or 'Townhouse'
2. **Apartment**: Includes properties labeled as 'Apartment' or 'Condominium'
3. **Other**: Includes properties labeled as 'Bed & Breakfast', 'Boat', 'Bungalow', 'Cabin', 'Camper/RV', 'Chalet', 'Dorm', 'Earth House', 'Hut', 'Igloo', 'Loft', 'Other', 'Tent', 'Treehouse', Villa' and 'Yurt.

Within the long-term rental data, property types fall into eight categories.  Like the AirBNB data we group these into three categories matching as best as possible to the AirBNB categories:

1. **House**:  Includes properties labeled 'Duplex', 'House', 'Terrace' or 'Townhouse'
2. **Apartment**: Includes properties labeled as 'Unit' or 'Studio'
3. **Other**: Includes properties labeled as 'Semi' and 'Villa'

Part of the difficulty in perfectly mapping property types from these two datasets is due to the fact that the long-term data uses Australian terms while the short-term data conforms to North American lexicon.  This is most apparent when talking about apartment dwellings.  In Australia these are referred to as Units or Studios, while in North American they are referred to as Apartments or Condominiums (depending on ownership structure)^[Also note that 'Apartment-type' ownership is very rare in Australia.  Nearly every multi-family dwelling is owned as a condominium.  See @berry2000investment for more information on the overall structure of the Australian market]. Additionally, Terrace homes are very common in many of the older suburbs of Melbourne and would be considered Rowhouses or, likely, Townhomes in the North American context.

The **Other** category only makes up a very small percentage of the properties in both the short and long term market; especially so in the long-term market (Figure 1) .  Additionally, a large proportion of the **Others** in the short-term data are Bed & Breakfast units, a use which doesn't fit with our research question. As a result, we remove the 'Others' from both the short- and the long-term rental data.

```{r load_beforePT, message=FALSE, warning=FALSE, cache=TRUE, comment=FALSE, echo=FALSE, results='hide'}
  data.path <- 'c:/dropbox/research/airBNB/data/'
  load(paste0(data.path, '/analyzed/analyzed_workspace.RData'))
```

```{r load_ptplot, message=FALSE, warning=FALSE, cache=TRUE, comment=FALSE, echo=FALSE, results='hide', fig.height=3, fig.cap='Tenure Type by Property Type', fig.lp='typecomp'}

  load(paste0(fig.path, 'ltrtypes.RData'))
  load(paste0(fig.path, 'strtypes.RData'))

  ggMultiPlots(str.types, ltr.types, cols=2)
  
```

Next, the short-term rentals properties can be listed as one of three types relating to the extent of the property which is able to be booked:

1. **Entire Home/Apt**: The entire home or apartment is available
2. **Private Room**: One room within a house or apartment is available
3. **Shared Room**: A bed within a room shared by another occupant(s) is available

As long-term rentals do not offer **Private Room** or **Shared Room** options^[Our dataset does not include rooming houses, purpose built student accommodations and other types of rentals which may offer private or shared room options] and our purpose here is a comparison of long- and short-term returns, we remove all short-term properties that do not lease the entire home or apartment. Unfortunately, this filter does remove about 40% of the short-term data (Figure 2), however, given the research question it is unavoidable.

```{r ltype_comp, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, fig.height=2.5, fig.width=3.75, fig.cap='Listing Type of Short Term Rentals'}


 load(paste0(fig.path, 'listtypes.RData'))

 list.types

```

#### Structural Characteristics 

We now turn to filtering by structural characteristics. As the Airbnb data provides limited information on the structural characteristics, the only overlapping fields that we have between the two datasets are bedrooms and bathrooms. Additionally, as the markets and structural components of houses and apartments differ, we bifurcate our analysis between the two property types from this point forward. 

We begin by removing all properties that are missing bedroom or bathroom information or have a likely data error in this fields (more than 14 bedrooms or 10 bathrooms). For both short and long term rentals we find that most apartments have fewer than four bedrooms and houses fewer than five.  For both property types, the vast majority of properties have three or fewer bathrooms. We filter the data accordingly. Additionally, note that short term rentals consider studio apartments to be 0 bed units while the long term data considers these as 1 bedroom. We convert the 0 bed short term units to 1 bed units. A few short term properties indicate no baths, which is likely an error so we remove these observations from the data. Also, here we notice that the short term data gives bathrooms in halves, while the long term data presents baths in whole numbers.  We round the baths in the short term data up to whole integers.

In addition to looking at bedrooms and bathrooms as separate dimensions of properties, they can also be considered together.  In other words, it is very common in the industry (and when looking for a short-term rental) to specify searches by bed/bath combination for obvious reasons -- 1 bedroom with 3 baths is inefficient and 4 bedrooms with 1 bath is uncomfortable.  As a result we examine the combinations of the two (Figure 3). From this analysis we see that there are 6 combinations that make up most all properties:  1bed/1bath, 2/1, 2/2, 3/1, 3/2 and 4/2. We filter the data to these combinations. 

```{r bb_comp, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, fig.height=4, fig.width=5.75, fig.cap='Bed/Baths by Property Type'}


 load(paste0(fig.path, 'ltrbb.RData'))
 load(paste0(fig.path, 'strbb.RData'))
 
 stbb <- stbb + ylab('# of Obs')
 ltbb <- ltbb + ylab('# of Obs')
 
 ggMultiPlots(stbb, ltbb)
  
```

### Rates and Rental values

Next we filter by the prices -- the nightly rates and weekly rental values. The long term rental data have a single price observation, the asking rent at closing.  For the short term data, there are two indicators of nightly rate.  First, each property has a average nightly rate figure that takes the average of the rates for all booked nights over the October 2014 to August 2016 period. Second, there is a rate for each day in the daily observations. For the vast majority of properties, the daily rates change very little, but for a select few there is a wide variation in the daily rates, often somewhat randomly. In cases where a small number of reservations have been made, the average nightly rate can deviate widely from usual rate due to outliers on the high end. To remedy this, we create a new property-level variable that is the median nightly rate of successful bookings over the September 1, 2015 to August 31, 2016 period.

We start by removing property level observations missing price data. Plotting the distribution of the rates and rents we see that most long term weekly rentals are greater than \$200 and less than \$1,000 per week while most median nightly rates are greater than \$50 and less than \$500 per night (Figure 4). We filter the property level data accordingly.  We also filter the short term daily data to include only those observations with nightly rates between \$50 and \$500.

```{r ratesrents, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, fig.height=4, fig.width=5.75, fig.cap='Melbourne Suburb Classifications'}


 load(paste0(fig.path, 'rateplot.RData'))
 load(paste0(fig.path, 'rentplot.RData'))

 ggMultiPlots(rate.plot, ltr.plot, cols=2)
  
```

<!-- #### Observed Revenue -->

<!-- Our final data filtering considers the observed revenue of the properties. For the long-term rentals the observed revenue is the weekly rate multiplied by 52 minus the number of weeks the property was on the market.  In a number of extreme cases, time on market was extremely long and the total observed revenues very low. To remove these outlying observations, we filter out any properties with less than $5,000 in revenue. -->

<!-- For the short-term data, we first calculate the total number of bookings, the occupancy rate, blocked rate and proportion of the year over which the property was listed.  Using this data we sum up the total revenue from bookings. For properties that were not listed on the site for the entire year, we impute their annual revenue by dividing the actual revenue by the proportion of the year in which they were listed.^[Properties listed in summer, but not winter may be over imputed and vice versa, however, we expect the biases to cancel out over the entire sample.] Properties with no reserved bookings, and, therefore, no revenue, are removed. -->

<!-- Looking at the distribution of the revenues from each tenure type, we see that those for the long-term rentals are for more consistent, with the vast majority totaling between \$12,000 to \$30,000 per year (Figure 10).  Short term revenues range from neglible to highs of greater than \$80,000.  While this simple analysis does not control for location and product differences, it does highlight the fact that some Airbnb owners are likely making greater revenues than in long-term rentals, but that most are likely not. Also, the high number of very low yield Airbnb properties also suggests that some owners may not be focused on maximizing revenue, but instead prefer the flexibility of short-term leasing. -->

<!-- ```{r load_workspace2, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, cache=FALSE} -->

<!--  load(paste0(data.path, '/analyzed/abb_objs.RData')) -->

<!-- ``` -->

<!-- ```{r revdensity.plot, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=6,fig.height=3.5, fig.cap='Comparison of Revenue Distributions'} -->

<!--   abbr <- abb.revs[,c('bedbath', 'revenue')] -->
<!--   abbr$tenure <- 'short-term' -->
<!--   ltrr <- ltr.revs[,c('bedbath', 'revenue')] -->
<!--   ltrr$tenure <- 'long-term' -->
<!--   allr <- rbind(abbr, ltrr) -->

<!--   ggplot(allr, aes(x=revenue, fill=tenure, color=tenure)) + -->
<!--     geom_density(alpha=.5) + -->
<!--     scale_fill_manual(values=c(ltr.col, abb.col)) + -->
<!--     scale_color_manual(values=c(ltr.col, abb.col)) + -->
<!--     ggtitle('Annual Revenues by Tenure Type') + -->
<!--     xlab('Annual Revenue') + -->
<!--     scale_x_continuous(breaks=c(seq(0,60000, by=20000)), -->
<!--                        labels=c('$0', '$20,000', '$40,000', '$60,000'))+ -->
<!--     theme(legend.position='bottom', -->
<!--           legend.title = element_blank(), -->
<!--           plot.title = element_text(hjust = 0.5), -->
<!--           axis.text.y = element_blank(), -->
<!--           axis.ticks.y = element_blank(), -->
<!--           axis.title.y=element_blank()) + -->
<!--     coord_cartesian(xlim=c(0, 70000)) -->

<!-- ``` -->

### Location and Submarkets

Location is an important determinant in both the short- and long-term rental markets. At the broadest scale, the long-term Melbourne residential market is usually discussed in terms of Inner, Middle and Outer suburbs.  In general, prices are highest in the inner suburbs and lowest in the outer, with a number of exceptions in the high end neighborhoods in the east and southeastern areas of the middle suburbs.  Suburbs in Melbourne are much smaller than their North American counterparts. The specific suburb of a property is the second, or finer scale, at which the market operates. Figure 5 shows the three large submarkets (Inner, Middle and Outer) and the individual suburbs boundaries.

```{r locmap, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, fig.height=4, fig.width=4, fig.cap='Melbourne Submarkets'}

 load(paste0(fig.path, 'suburbMap.RData'))

 loc.map 
 
```

It is likely that the short-term rental market follows a similar market hierarchy, however, we can imagine that a few different spatial features can influence the short-term market.  While there are many possible additional features, we consider two to be the most likely to influence rates or occupancy: 1) Proximity to beaches; and 2) Proximity to key tourist activities and events. Using the three broad submarkets from the long-term market (inner, middle and outer) we have created a five sub-market system as a starting point for spatially analyzing short and long term rentals in Melbourne: 

1. Rural (Outer suburbs, not Beach)
2. Suburban (Middle suburbs, not Beach)
3. City (Inner Suburbs, not Beach, not Core)
4. City-Core (Select inner suburbs with tourist activities and near CBD)
5. Beach (Properties within 500m of Port Phillip Bay east of Yarra River)

We select the 18 suburbs below to represent the 'core' of the city (Table 1).  The majority of tourist destinations and major events such as the Australian Open and the Grand Prix are located in these suburbs.  All are well serviced by public transportation and possess abundant amenities for tourists.

```{r, results="asis", comment=FALSE, message=FALSE, echo=FALSE}

 c.subs <- data.frame(A=core.suburbs[1:6],
                      B=core.suburbs[7:12],
                      C=core.suburbs[13:18])
 names(c.subs) <- NULL
 rownames(c.subs) <- NULL

 print(xtable(c.subs, caption = 'List of Core Suburbs\n'),
       include.rownames=F,
       caption.placement="top",
       hline.after=c(0,6))


```

We assign these designations by: 1) Adding suburb designations to the properties; 2) Assigning submarkets 1-4 based on suburb location; and 3) Indicating proximity to beach and labeling as 'Beach' submarket.  This process is repeated for both the short and long term data. Finally, we also remove any observations that fall outside of outer suburbs (the extent of the suburbs). The location of the short and long term rentals, colored by submarket, are shown in Figure 6.

Before assigning submarkets to the long term property we must remedy the fact that about 5% of the long term rental observations have missing latitude and longitude values.  Half of these do, however, have a latitude and longitudeg value for the centroid of the street that the property faces.  To retain as many observations as possible in these intial steps we apply the street centroid value to the observations with missing lat/long values.  For those without either set of lat/longs we remove them from the dataset at this point. After making this correction we then assign one of the five submarket designations based on the process described above.

```{r loc.plots, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=5.5, fig.height=9, fig.cap='Location Maps'}

  load(paste0(fig.path, 'ltrMap.RData'))
  load(paste0(fig.path, 'strMap.RData'))

  ggMultiPlots(str.map, ltr.map, cols=1)

```

To better visualize the relative frequency of the submarkets in the two tenure types we have broken down the data by submarket by property type by tenure type. A number of interesting observations are illustrated here (Figure 7). First, the vast majority of Airbnb units are apartments in city-core or city locations. As expected and as shown by the map, the long term properties are more evenly spread across the metro region than the short term properties.  Long term houses are more peripheral than apartments, a result of urban development patterns and land economics. Beach properties are, relatively speaking, over-represented in the short term market, especially in apartments, while suburban and rural apartments, somewhat common in long term markets, are nearly absent in the short-term market. Overall, the suburban and rural Airbnb markets are quite thin, later results focusing on these markets will need to be examined in light of this small sample size.

```{r loc.bars, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=8.5, fig.height=4, fig.cap='Location Maps'}

  load(paste0(fig.path, 'ltrsubplot.RData'))
  load(paste0(fig.path, 'strsubplot.RData'))
  ltrtype.map <- ltrtype.map + ylab("")

  ggMultiPlots(strtype.map, ltrtype.map, cols=2)

```


<!-- ### Summary of Data Cleaning -->

<!-- We have applied nine separate filters to our two datasets.  In the name of transparency and to highlight the bounds of our research questions it is useful to review this filters (Table 2). There are more than 20,000 properties that are or have been listed on the Airbnb portal in Melbourne during the 2014 to 2016 period.  Filtering those not listed during the study period removes 1/3 of all properties.  Further removing by property type and listing type (not entire home) elimates another 50% of the remaining properties.  From these 6,600 properties structural characteristics, locational issues and outlying nightly rates remove another 800 from consideration.  Finally, 1,400 of these have never been booked and have \$0 revenue, requiring removal.  From over 20,000, we have 4,446 Airbnb properties that meet the requirements of our research question. -->

<!-- For the long term data, over half of the observations fall outside of our considered time frame, reducing the sample size down to around 130,000 from 290,000.  Further consideration regarding property types, structural characteristics, location, etc. remove another 10,000 observations, leaving 122,051 in the final dataset. -->

<!-- ```{r, results="asis", comment=FALSE, message=FALSE, echo=FALSE, fig.cap='Data Cleaning Summary'} -->

<!-- cc <- clean.count[,c('operation', 'abb', 'ltr')] -->
<!-- abb.count <- cumsum(c(cc$abb[1], -cc$abb[-1])) -->
<!-- ltr.count <- cumsum(c(cc$ltr[1], -cc$ltr[-1])) -->
<!-- abb.cut <- c(0, abb.count[-length(abb.count)] - abb.count[-1]) -->
<!-- abb.per <- c(0, abb.cut[-1] / abb.count[-length(abb.count)]) -->
<!-- ltr.cut <- c(0, ltr.count[-length(ltr.count)] - ltr.count[-1]) -->
<!-- ltr.per <- c(0, ltr.cut[-1] / ltr.count[-length(ltr.count)]) -->
<!-- clean.table <- data.frame(field=cc$operation, -->
<!--                           abbcount=abb.count, -->
<!--                           abbcut=abb.cut, -->
<!--                           abbper=abb.per, -->
<!--                           ltrcount=ltr.count, -->
<!--                           ltrcut=ltr.cut, -->
<!--                           ltrper=ltr.per) -->

<!-- rownames(clean.table) <- NULL -->
<!-- clean.table$field <- c('Initial Count', 'Time', 'Property Type', -->
<!--                        'Listing Type', 'Bedrooms', 'Baths', 'Bed/Bath Combo', -->
<!--                        'Rates & Rents', 'Location', 'Revenue') -->
<!-- clean.table$abbper <- 100*round(clean.table$abbper, 2) -->
<!-- clean.table$ltrper <- 100*round(clean.table$ltrper, 2) -->
<!-- names(clean.table) <- c('Filter', 'Short Term', 'Cut', '% Cut', -->
<!--                         'Long Term', 'Cut', '% Cut') -->
<!-- options(xtable.comment = FALSE) -->
<!-- print(xtable(clean.table, caption='Data Cleaning Summary\n', -->
<!--        digits=0), -->
<!--        include.rownames=FALSE, -->
<!--        caption.placement='top') -->

<!-- ``` -->

<!-- ## Comparison -->

<!-- In this section we compare returns between long and short term rental options.  As mentioned above, there are no direct matches between the datasets. In other words, we have not directly observed a property which has been in a long-term rental followed by short-term or vice versa.  We will, instead, make the comparisons between the two by imputing likely short-term (long-term) rental revenues for the long-term (short-term) data and then comparing the observed revenues to the imputed revenues of the unobserved tenure type. -->

<!-- We make our imputations via hedonic price models, one for long-term rentals and one for short-term rentals.  We model the long-term market with the long-term data and then use this model to predict or impute the likely long-term rental for the short term properties.  We do the same for the long-term properties using a modeled derived from the short-term data. -->

<!-- For the long-term market, we use the following model specification: -->

<!-- \begin{equation} -->
<!--   \label{eq:test} -->
<!-- ln(RentLT_{i}) = \beta_{1} BedBath_{i} + \gamma_{2} Suburb_{i} + \tau_{3} Time_{i} + \epsilon_{i} -->
<!-- \end{equation} -->

<!-- where $ln(RentLT_{i})$ is the natural log of the weekly rent, *BedBath* is the bedroom/bathroom combination, *Suburb* is a fixed effect for the location of the property, *Time* is the month of rental and $\epsilon$ a random error term. -->

<!-- For the short-term market the specification is similar, minus the time correction.  We eliminate this control because the short term rates are based on the medium value over the entire period and not monthly. This specification is: -->

<!-- \begin{equation} -->
<!-- ln(RentST_{i}) = \beta_{1} BedBath_{i} + \gamma_{2} Suburb_{i} + \epsilon_{i} -->
<!-- \end{equation} -->

<!-- where *$ ln(RentLT_{i})$* is the natural log of the weekly rent, *BedBath* is the bedroom/bathroom combination, *Suburb* is a fixed effect for the location of the property and $\epsilon$ a random error term. -->


<!-- ## Results -->

<!-- We will begin our analysis by making a comparison of the short and long-term markets at a metro-level.  What we mean by this is that we will treat the entire region as the functional market and will impute rates and rents with a model using all data.  It is likely that different submarkets operate differently, a hypothesis investigated after the full metropolitan region. -->

<!-- We begin by imputing the likely long-term rents for the short-term observations, and vice versa at a metro level. A comparison between the actual (observed) and imputed rates and rents shows that, in general, the long-term properties are composed of properties that have a higher nightly rate than the short term observations (left-panel) (Figure 11).  This is likely due to the fact that there are more houses (compared to apartments) in the long-term data than in the short-term data. Interestingly, the observed (long-term) and imputed (short term) weekly rental values are somewhat similar, suggesting that the higher rent, city-core apartments push the distribution of the imputed short-term values upwards. -->

<!-- For each of the observed values we also create a fitted value for that observation in the model.  This may help to reduce the influence of outlying observed values in our model.  To be clear, each observation will now have three estimates of rates and rents.  For the short-term properties they will have: 1) an actual nightly rate; 2) an estimated nightly rate; and 3) an estimated long-term (weekly) rent.  The long-term properties will have: 1) an actual long-term (weekly) rent; 2) an estimated long-term (weekly) rent; and 3) an estimated nightly rate. -->

<!-- ```{r pred.glob.imp, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=6,fig.height=3.5} -->

<!--   abbir <- abb.revs[,c('bedbath', 'imp.rent', 'med.rate')] -->
<!--   abbir$tenure.rate <- 'short-term (observed)' -->
<!--   abbir$tenure.rent <- 'short-term (imputed)' -->

<!--   ltrir <- ltr.revs[,c('bedbath', 'event.price', 'imp.rate')] -->
<!--   ltrir$tenure.rate <- 'long-term (imputed)' -->
<!--   ltrir$tenure.rent <- 'long-term (observed)' -->

<!--   names(abbir) <- names(ltrir) <- c('bedbath', 'rent', 'rate', 'tenure.rate', 'tenure.rent') -->
<!--   allir <- rbind(abbir, ltrir) -->

<!-- ``` -->

<!-- ```{r imp.rate.plot, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=6,fig.height=3.5} -->

<!--   irate.plot <- ggplot(allir, aes(x=rate, fill=tenure.rate, color=tenure.rate)) + -->
<!--     geom_density(alpha=.5) + -->
<!--     scale_fill_manual(values=c(ltr.col, abb.col)) + -->
<!--     scale_color_manual(values=c(ltr.col, abb.col)) + -->
<!--     ggtitle('Nightly Rates by Tenure Type') + -->
<!--     xlab('Nightly Rate') + -->
<!--     scale_x_continuous(breaks=c(seq(0,600, by=200)), -->
<!--                        labels=c('$0', '$200', '$400', '$600'))+ -->
<!--     theme(legend.position='bottom', -->
<!--           legend.title = element_blank(), -->
<!--           plot.title = element_text(hjust = 0.5), -->
<!--           axis.text.y = element_blank(), -->
<!--           axis.ticks.y = element_blank(), -->
<!--           axis.title.y=element_blank()) + -->
<!--     coord_cartesian(xlim=c(0, 625)) -->

<!-- ``` -->

<!-- ```{r imp.rent.plot, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=6,fig.height=3.5} -->

<!--   irent.plot <- ggplot(allir, aes(x=rent, fill=tenure.rent, color=tenure.rent)) + -->
<!--     geom_density(alpha=.5) + -->
<!--     scale_fill_manual(values=c(ltr.col, abb.col)) + -->
<!--     scale_color_manual(values=c(ltr.col, abb.col)) + -->
<!--     ggtitle('Weekly Rents by Tenure Type') + -->
<!--     xlab('Weekly Rents') + -->
<!--     scale_x_continuous(breaks=c(seq(0,1000, by=200)), -->
<!--                        labels=c('$0', '$200', '$400', '$600', '$800', '$1,000'))+ -->
<!--     theme(legend.position='bottom', -->
<!--           legend.title = element_blank(), -->
<!--           plot.title = element_text(hjust = 0.5), -->
<!--           axis.text.y = element_blank(), -->
<!--           axis.ticks.y = element_blank(), -->
<!--           axis.title.y=element_blank()) + -->
<!--     coord_cartesian(xlim=c(150, 1000)) -->

<!-- ``` -->

<!-- ```{r imp.rr.plot, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=9,fig.height=3.5, fig.cap='Rates and Rents Comparison'} -->

<!--  ggMultiPlots(irate.plot, irent.plot, cols=2) -->

<!-- ``` -->

<!-- The question of whether or not the imputations are accurate remains.  Table 3 shows simple diagnostics from the two imputation models.  From this we see that the short-term model performs rather poorly, with a standard error of 34.6\% and over half of the residuals greater than 20\%.  The long term model, on the other hand, performs quite well, with a standard error under 16\% and less than 19\% of fitted errors greater than 20\%.  Given the small amount of independent variable available to the models, the long term market appear much more predictable than the short term market.  This could be due to omitted variables in the short term market or a greater level of variability in the price setting (which is done by individuals) as opposed to the long-term market in which most owners get help with leasing from an agency. -->

<!-- ```{r, results="asis", echo=FALSE} -->

<!-- abbs <- c(abb.sum$r.squared, abb.sum$adj.r.squared, -->
<!--           abb.sum$sigma, -->
<!--           length(which(abs(abb.sum$residuals) > .2)) / length(abb.sum$residuals)) -->
<!-- ltrs <- c(ltr.sum$r.squared, ltr.sum$adj.r.squared, -->
<!--           ltr.sum$sigma, -->
<!--           length(which(abs(ltr.sum$residuals) > .2)) / length(ltr.sum$residuals)) -->

<!-- summ.table <- data.frame(shortterm=round(abbs, 3), -->
<!--                          longterm=round(ltrs, 3)) -->
<!-- rownames(summ.table) <- c('r-squared', 'adj r-squared', -->
<!--                           'sigma', '% fit errors > 20%') -->
<!-- names(summ.table) <- c('Short Term', 'Long Term') -->

<!-- s.table <- xtable(summ.table, -->
<!--                   caption='Imputation Model Diagnostics\n', -->
<!--                   digits=3) -->
<!-- align(s.table) <- c('lcc') -->
<!-- print(s.table,  -->
<!--        include.rownames=TRUE, -->
<!--                   caption.placement='top') -->


<!-- ``` -->

<!-- Next, to create an estimated short-term revenue for the long-term properties we have to create an estimate of the occupancy rate. For the estimated long-term rentals we have to create an estimate of the search time (days on market).  To derive an occupancy rate, we use the weighted median value^[Weighted by the total number of days the property was offered on Airbnb over the year.] of the market, the entire metro region in this case, which is about 37%^[We attempted to use a logistic regression model to predict occupancy but were unsucessful to develop an appropriate approach.  Work continues toward this end.].  The range of occupancy varies widely from essentially 0 to 100%, with a standard deviation of 27%. -->

<!-- For the long term search time, or days on market, we also use the median value of the observations. The median days on market is 24, with a standard deviation of 37. About 60\% of long term rental spent one month or less on the market and 85\% spent less than two months searching for a tenant. -->

<!-- Using these estimated occupancy rates and days on market, we have used equations 3 and 4 to produce estimates of long-term and short-term revenues for the short and long term observations, respectively. We do these for the observed and the estimated rates and rents for each tenure type. As we have imputed and actual observations for one of each dataset in the comparison, this creates 4 different comparisons, roughly broken down as shown in Figure 11. -->

<!-- ```{r, results="asis", echo=FALSE, fig.width=4,fig.height=4, fig.cap='Revenue Comparison Types'} -->

<!--   explain.plot <- blank.plot + -->
<!--     geom_text(data=ann.df, -->
<!--               label=ann.df$lab, -->
<!--               size=2.8) + -->
<!--     geom_text(data=num.df, -->
<!--               label=num.df$count, -->
<!--               size=9.5) + -->
<!--     coord_cartesian(ylim=c(-.5,1)) + -->
<!--     ggtitle('Explanation of \n Revenue Comparisons') + -->
<!--     theme(legend.position='bottom', -->
<!--           legend.title = element_blank(), -->
<!--           plot.title = element_text(hjust = 0.5)) -->

<!-- ``` -->

<!-- ```{r, results="asis", echo=FALSE} -->

<!--   one.df <- data.frame(x=rep(0, 4), -->
<!--                        y=rep(0, 4), -->
<!--                        lab=c('31%', '87%', '54%', '79%'), -->
<!--                        est=c('Actual Rates & Rents', 'Actual Rates & Rents', -->
<!--                              'Imp. Rates & Rents', 'Imp. Rates & Rents'), -->
<!--                        data=c('Airbnb', 'Long-Term', 'Airbnb', 'Long-Term')) -->

<!--   sizes <- c(.31, .87, .54, .79) -->

<!--   one.plot <- blank.plot + -->
<!--     geom_text(data=one.df, -->
<!--               label=one.df$lab, -->
<!--               size=18*sizes) + -->
<!--     ggtitle('% of Properties where \n Airbnb is more profitable') + -->
<!--     theme(legend.position='bottom', -->
<!--           legend.title = element_blank(), -->
<!--           plot.title = element_text(hjust = 0.5)) -->

<!-- ``` -->

<!-- In other words, the left hand panels (left side of Figure 12) are comparisons made on Airbnb properties, where the top is the actual Airbnb revenue and the bottom is the estimated (fitted) with the imputed rates and median-based occupancy rates.  In the right hand panels we have the comparison on the long-term properties, again with top being of the actual long term revenue and the bottom with the estimated (fitted) rents and median based days on market. -->

<!-- The right hand side of Figure 12 shows the percent of properties in which a short-term rental would bring in more revenue over the coures of the year, across these four scenarios.  The variation between these estimates is significant and troublesome.  However, given our difficulty in accurately estimating short term nightly rates (see Table 3) and the fact that occupancy rates are highly variable for short-term rentals, these variations are not surprising.  As we have a difficult time estimating short-term market performance, both measures in the right hand column are immediately suspicious. Additionally, the estimates in the lower left also depend on estimate nightly rate, causing their accuracy to be questioned. -->

<!-- ```{r 22r.plot, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=7.5,fig.height=3.5, fig.align='center', fig.cap='Summary of Comparison Types'} -->

<!--  ggMultiPlots(explain.plot, one.plot, cols=2) -->

<!-- ``` -->

<!-- In short, only in the upper left hand panel are we using the actual nightly rates and the actual occupancy rates from the short term market.  We do estimate the long-term rates and days on market here, but as shown above, the models for those figures are much more robust. The plusses and minus of each comparison quadrant are shown in Figure 13.  As a result of these issues, we consider only those comparisons in quadrant 1 to be legitimate and use these going forward.  This means that we will only be considering comparisons of actual Airbnb revenues to their estimated long-term rental revenues.  -->

<!-- ```{r reason.plot, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=4.5, fig.height=3.5, fig.align='center', fig.cap='Comparison Types: Strengths and Weaknesses'} -->

<!--   reason.plot <- blank.plot + -->
<!--     geom_rect(data = mrkt.table, aes(fill=good), -->
<!--               xmin = -Inf,xmax = Inf, -->
<!--               ymin = -Inf,ymax = Inf, alpha = 0.4) + -->
<!--     geom_text(data=reason.df, -->
<!--               label=reason.df$lab, -->
<!--               size=3) + -->
<!--     geom_text(data=num.df, -->
<!--               label=num.df$count, -->
<!--               size=9.5) + -->
<!--     coord_cartesian(ylim=c(-.5,1)) -->

<!--   reason.plot -->

<!-- ``` -->

<!-- ### Occupancy Rates -->

<!-- Occupancy rates are one of the most important components in determining the profitability of short-term rentals. Unfortunately, these are also the most difficult factor to estimate or impute. Below we shows (left-panel, Figure 14) the percentage of properties for which short-term rental is more profitable versus the occupancy rate.  We see that likelihood of short-term rental being a better option increases sharply with occupancy rate up until about 75\% occupancy.  For occupancy rates less than 50\%, long-term rental is more likely to be the better option.  The decrease in profitabilty after 75\% occupancy is likely due to: 1) small sample sizes in this tail of the data; and 2) that lowering overall nightly rates too low simply to boost occupancy can result in lower overall revenue. -->

<!-- As the distribution of occupancy rates is not uniform (Figure 14, right panel), the initial plot may be misleading as to the frequency of short-term rentals being more profitable than long term. In Figure 15, we transform the x-axis to represent the quantile of the full distribution of occupancy rates. By doing so, we see that, in general, properties need to be in the upper 1/3 of occupancy rates across the entire market in order to have a equal likelihood of being more profitable as a short-term rental than as a more traditional long term one. -->

<!-- ```{r message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE} -->

<!--   rawocc.pplot <- makePrefPlot(abb.revs, -->
<!--                                x.field='occ.rate', -->
<!--                                y.field='abb.act', -->
<!--                                     smooth=TRUE, -->
<!--                                smooth.span=.75) -->
<!--   rawocc.pplot <- rawocc.pplot + -->
<!--     xlab('\nOccupancy Rate') + -->
<!--     ylab('\n% of Properties where \nAirbnb is more profitable') + -->
<!--     scale_x_continuous(breaks=seq(0, 100, by=25), -->
<!--                        labels=c('0%', '25%', '50%', '75%', '100%')) + -->
<!--     scale_y_continuous(breaks=seq(0, 1, by=.25), -->
<!--                        labels=c('0%', '25%', '50%', '75%', '100%')) + -->
<!--     scale_color_manual(values=sm.col) + -->
<!--     ggtitle('Tenure Profitability vs \n Occupancy Rate') + -->
<!--     theme(legend.position='none', -->
<!--           plot.title = element_text(hjust = 0.5)) + -->
<!--     coord_cartesian(ylim=c(0,1)) + -->
<!--     geom_hline(aes(yintercept=.5)) -->


<!-- ``` -->

<!-- ```{r message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE} -->

<!--   dens.plot <- ggplot(abb.revs, aes(x=occ.rate)) + -->
<!--     geom_density(alpha=.3, fill=1) + -->
<!--     scale_color_manual(values=sm.col) + -->
<!--     scale_fill_manual(values=sm.col) + -->
<!--     scale_x_continuous(breaks=seq(0, 1, by=.25), -->
<!--                        labels=c('0%', '25%', '50%', '75%', '100%')) + -->

<!--     ylab('\nFrequency of Occurance') + -->
<!--     xlab('\nOccupancy Rate') + -->
<!--     ggtitle('Distribution of \n Occupancy Rates') + -->
<!--     theme(legend.position='none', -->
<!--           plot.title = element_text(hjust = 0.5), -->
<!--           axis.line=element_blank(), -->
<!--           axis.text.y=element_blank(), -->
<!--           axis.ticks.y=element_blank()) -->
<!-- ``` -->

<!-- ```{r message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=7, fig.height=3.5, fig.cap='Probability of Short Term Pref. vs Occ. Rate'} -->

<!--   ggMultiPlots(rawocc.pplot, dens.plot, cols=2) -->

<!-- ``` -->

<!-- ```{r message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE} -->

<!--   abb.revs$occ.qtl <- makeWtdQtl(abb.revs$occ.rate, -->
<!--                                       return.type='rank') -->
<!--   abb.revs$rate.qtl <- makeWtdQtl(abb.revs$med.rate, -->
<!--                                   return.type='rank') -->

<!--   ## Make quartile location plot -->

<!--   qtlocc.pplot <- makePrefPlot(abb.revs, -->
<!--                                x.field='occ.qtl', -->
<!--                                y.field='abb.act', -->
<!--                                smooth=TRUE, -->
<!--                                smooth.span=.55) -->
<!--   qtlocc.pplot <- qtlocc.pplot + -->
<!--     xlab('\nQuantile of Occupancy Rate') + -->
<!--     ylab('\n% of Properties where \n Airbnb is more profitable') + -->
<!--     scale_x_continuous(breaks=seq(0, 100, by=25), -->
<!--                        labels=c('0', '25th', '50th', '75th', '100th')) + -->
<!--     scale_y_continuous(breaks=seq(0, 1, by=.25), -->
<!--                        labels=c('0%', '25%', '50%', '75%', '100%')) + -->
<!--     scale_color_manual(values=sm.col) + -->
<!--     theme(legend.position='none') + -->
<!--     coord_cartesian(ylim=c(0,1))+ -->
<!--     geom_hline(aes(yintercept=.5)) -->

<!-- ``` -->

<!-- ```{r message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=4, fig.height=3, fig.align='center', fig.cap='Probability of Short Term Pref. vs Occ. Rate Quantile'} -->

<!--   qtlocc.pplot -->

<!-- ``` -->

<!-- Along with occupancy rates, the other important consideration is the nightly rate.  High occupancy rates with low nightly rates may not be any better than lower occupancy rate with a higher nightly rate.  In Figure 16 (left panel) we plot nightly rates versus occupancy rates.  We then divide the area of the scatter plot into 400 equal areas (20 x 20) and color each based on which is more common in the area, short term or long term revenue.  The transparency is related to the total count of actual market properties falling into that area. -->

<!-- The general trend is what is expected, i.e. that higher rates and occupancies, ceteris paribus, means more profitability for short term rentals.  This relationship is not a perfectly linear because the long-term rental revenue varies greatly based on the characteristics of the property.  While the left panel of Figure 16 helps show the general pattern is does leave many indeterminate areas.  In order to better delimit the situations where short-term rentals are more profitable we turn to a support vector machine (SVM) algorithm. -->

<!-- An SVM can be used as a classifier to delineate the boundaries between two (or more) classes of observations.  In this case, we delineate the boundary (right panel) between situations where long-term rentals are more profitable (red) and where short term are (green).  In order to avoid an overly broad straight line delineation we use a polynomial fitting technique. While this process does provide more flexible fit in the area of high data density, it also has the side effect of creating somewhat counter intuitive results in areas of low data (such as the far upper right corner.)  Overall, from the SVM model we see that the short-term preference begins at about 30\% occupancy rate, regardless of nightly rate.  As the occupancy rate increases, the required nighly rate decreases, down to near $100 once occupancies hit 65\% or greater. -->

<!-- ```{r, fig.width=8, fig.height=4, echo=FALSE, fig.align='center', fig.cap='Analysis of Raw Rates: Metro Level'} -->

<!--  ggMultiPlots(rate.hm, rate.hm.svm, cols=2) -->

<!-- ``` -->

<!-- Both the occupancy rates and the nightly rates are not uniformly distributed and therefore the plots in Figure 16 may be misleading.  We convert both axis to quantiles and replot both in Figure 17.  Here we see a more even distibution of points across the figures.  Additionally, a more reasonable SVM breakdown occurs as well.  The right panel shows that in order for the likelihood of a property to be more profitable in short-term rental to be above 50\% the lowest occupancy rate it can have is 40\%, if the nightly rate is in the top 95th percentile of the market.  For example, if we drop the nightly rate to the median, the occupancy rate need to be at the 60th percentile. -->

<!-- ```{r, fig.width=8, fig.height=4, echo=FALSE, fig.cap='Analysis of Quantile Rates: Metro Level'} -->

<!--  ggMultiPlots(qtl.hm, qtl.hm.svm, cols=2) -->

<!-- ``` -->

<!-- ### Market Summary -->

<!-- Finally, we summarize the entire market by looking at three metrics: -->

<!-- 1. Actual: The percentage of properties in the market in which short-term (Airbnb) generates more revenue -->
<!-- 2. Fitted: The percentage of properties (when plotted by occupancy rate and nightly rate) that fall in the SVM area considered profitable -->
<!-- 3. SVM Area: The percentage of the total area in the SVM plot that represents a situation where the short-term rental is more profitable. -->

<!-- None of these metrics on their own is perfect.  The 'actual' measure responds directly to the sample of the current market and may not be as useful to a potential investor. The 'fitted' and 'svm area' metric minimize some bias from the sample, but suffer from forcing all observations into a binary condition (there are no 'gray' areas) and the 'svm area' metric may be biases by large areas of the scatterplot that have few actual observations.  This latter issue is partially limited by plotting by quantile, but, nonetheless still exists. -->

<!-- We calculate each for the raw rates as well as for the quantile situation.  These are show in Table 4.  Overall, 31.4\% of the observed Airbnb properties raised more actual revenue from short-term lending than the estimated long-term revenue that they would have gained.  Analyzing these by raw rates in the SVM, we find that a nearly identical 31.4\% fall within the contiguous area of short-term preference and this area takes up 40.6% of the total plotted area.  When we scale the plots by quantile, we find that 37.2\% of all observations fall in the short-term preference area and that this area takes up 37.8\% of the total plotted area.  As the difference between the fitted and the svm area is considerably smaller in the quantile approach, there is less bias from outlying, empty areas. -->

<!-- ```{r, results="asis", comment=FALSE,  echo=FALSE} -->

<!-- names(market.ratio) <- c('Property Type', 'Actual', 'Fitted', 'SVM Area') -->
<!-- print(xtable(market.ratio, caption='Metro Level: Short-Term Preference Analysis\n', -->
<!--        digits=3), -->
<!--        include.rownames=FALSE, -->
<!--       caption.placement='top') -->


<!-- ``` -->

<!-- ### Submarkets -->

<!-- A major concern about the above analysis is that the estimated rents, rates, days on market and occupancy rates are estimated globally.  As a result, there is the likelihood of significant bias within certain submarkets.  Below, we break down the results by type (apartment vs house), by geographic market (five geographic submarkets listed above) and by type and submarket.   -->

<!-- #### Property Type -->

<!-- First, we bifurcate our data into houses and apartments.  We then rerun the entire analysis described above for the house and the apartment markets separately. Table 5 and Figures 18 (houses) and 19 (apartments) illustrate the metrics for these two sub-markets.  The raw probability data shows that over 33% of apartments are more profitable in short term leasing as opposed to just 22% of houses.  Figures 18 and 19 show the quantile preference plots both in the raw and with the SVM categorization.  As is evident from these plots, the likely of short-term rental being more profitable is much greater in the apartment market than in the house market.  For apartments, a 60th rate and a 60th percentile occupancy rate has a better than 50% chance of being more profitable in a short term leasing strategy than a long term one.  In fact, so long as an apartment is in the top 1/4 of occupancy rates, it is more profitable in a short term strategy at almost any price point.  Houses, on the other hand, need either very high rates or very high occupancies to justify a short-term rental.   -->

<!-- The large differences between the `SVM Area' figures and the Actual and Fitted speaks to two points about the data.  First, the difference between the SVM Area and the Fitted number highlight the fact that even when transforming the axes into quantile, the data are still not uniformly distributed across the 2-dimensional space of the plot.  This means that not all areas of the plot have equal weights in the Fitted value.  Second, the Actual and SVM Area figures deviate because of the binary nature of the SVM process.  For example, in many of the short-term preference in the apartment SVM plot there are some long-term preference data points, but just fewer than 50% of total, whereas in the long-term preference areas on the left hand side of the plot nearly all of the points are long-term preference.  The results of the SVM process are the result of a sort of 'gerrymandering' where a clustering by one faction and a dispersal by another can create a result that is not representative of the underlying base population.  Despite this fact, we feel that the SVM Area values are useful for the owner in our scenario because they provide both a value and visual guide that present the possibilities of being more profitable with a short-term strategy, not just what the current distribution of such properties are.  -->


<!-- ```{r, results="asis", echo=FALSE, fig.cap='Preference Analysis: By Property Type'} -->

<!--  type.table <- rbind(apt$mrkt.score, house$mrkt.score) -->
<!--  type.table$property.type <- c('apartment', '', 'house', '') -->
<!--  type.table <- type.table[,c(5,1:4)] -->
<!--  names(type.table) <- c('Property Type', 'Actual', 'Fitted', 'SVM Area') -->
<!--  print(xtable(type.table, caption='Short-term Preference Analysis: By Property Type\n', -->
<!--        digits=3), -->
<!--        include.rownames=FALSE, hline.after=c(-1,0,2,4), -->
<!--        caption.placment='top') -->


<!-- ``` -->


<!-- ```{r, fig.width=8, fig.height=4, echo=FALSE, fig.align='center', , fig.cap='Analysis of Quantile Rates: Houses'} -->

<!-- ggMultiPlots(house$qtl.hm, house$qtl.hm.svm, cols=2) -->

<!-- ``` -->

<!-- ```{r, fig.width=8, fig.height=4, echo=FALSE, fig.align='center', , fig.cap='Analysis of Quantile Rates: Apartment'} -->

<!-- ggMultiPlots(apt$qtl.hm, apt$qtl.hm.svm, cols=2) -->

<!-- ``` -->

<!-- #### Geographic Submarkets -->

<!-- We break the data down further by geographic submarket and property type and compare the results (quantile) in Table 6.  In the Actual (observed) revenues, the probabilty of short-term preference is highest in the core area of the city and lowest in the city, beach and suburban areas.  Interestingly, the rural area has a higher probably than all other but the core.  This is likely explained by the fact that there are few rural properties and many of these are clustered near rural tourists activities like the Yarra Valley wine region which see a high demand for lodging (Figure 8).  -->

<!-- In the core and the city, apartments offer a higher probability than homes, the reverse is true in the suburban and rural areas. In the lower density areas it appears that tourists may prefer larger and more spacious dwellings compared to long-term renters who may experience long commutes from these fringe locations. -->

<!-- Looking at the 'fitted' and 'svm area' figures we see a trend that the svm values are higher than the actual observed probabilities for the core apartment and lower for nearly all of the rest of the categories.  This finding suggests that for all but the core apartments, those properties that are more profitable as short-term cluster tightly in the area (the combination of nightly rate and occupancy rate) that are preferable for a short-term strategy. In other words, there is well defined combination or situation in which these properties can be more profitable short-term than in a long-term lease.  The opposite is true of the city core apartments. Here, the area of the plots occupied by short term preference is larger than the total proporation, suggesting that a wide variety of combinations of nightly rates and occupancy rates will result in a short term preference. -->

<!-- ```{r, results='asis', echo=FALSE, message=FALSE, comment=FALSE, fig.cap='Quantile Market Summary; Property Type by Submarket'} -->

<!-- sm.ms <- rbind.fill(lapply(sm.results, function(x) x$mrkt.score[2,])) -->
<!-- sm.ms.apt <- rbind.fill(lapply(smt.results[1:5], function(x) x$mrkt.score[2,])) -->
<!-- sm.ms.house <- rbind.fill(lapply(smt.results[6:10], function(x) x$mrkt.score[2,])) -->

<!-- smt.table <- data.frame(sub.market=c('sub-market','beach', 'city', 'core', 'rural', -->
<!--                                      'suburban', 'all'), -->
<!--                         actual.all=c('all', round(sm.ms$actual[c(5,2,1,4,3)],2), -->
<!--                                      round(market.ratio[2,2],2)), -->
<!--                         actual.apt=c('apt', round(sm.ms.apt$actual, 2), -->
<!--                                      round(type.table[2,3],2)), -->
<!--                         actual.house=c('house', round(sm.ms.house$actual, 2), -->
<!--                                        round(type.table[4,3],2)), -->
<!--                         fitted.all=c('all',round(sm.ms$fitted[c(5,2,1,4,3)],2), -->
<!--                                      round(market.ratio[2,3],2)), -->
<!--                         fitted.apt=c('apt', round(sm.ms.apt$fitted,2), -->
<!--                                      round(type.table[2,4],2)), -->
<!--                         fitted.house=c('house', round(sm.ms.house$fitted,2), -->
<!--                                        round(type.table[4,4],2)), -->
<!--                         svmarea.all=c('all', round(sm.ms$svm[c(5,2,1,4,3)],2), -->
<!--                                       round(market.ratio[2,4],2)), -->
<!--                         svmarea.apt=c('apt', round(sm.ms.apt$svm,2), -->
<!--                                       round(type.table[2,5],2)), -->
<!--                         svmarea.house=c('house', round(sm.ms.house$svm,2), -->
<!--                                         round(type.table[4,5],2))) -->

<!-- smt.table <- smt.table[c(1, 4, 3, 2, 6, 5, 7),] -->
<!-- names(smt.table) <- c('', '', 'actual', '', '', 'fitted', '', '', -->
<!--                       'svm area', '') -->
<!-- smt.full <- xtable(smt.table, -->
<!--                    caption='Short-Term Preference Analysis: Disaggregated\n') -->
<!-- align(smt.full) <- 'll|rrr|rrr|rrr' -->
<!-- print(smt.full,  -->
<!--        include.rownames=FALSE,  -->
<!--        hline.after=c(-1,0,1,6,7,7), -->
<!--                    caption.placement='top') -->


<!-- ``` -->

<!-- ## Explaining Preference -->

<!-- The analyses above show that, overall, just less than 1 out of every 3 Airbnb'd properties was actually  more profitable as a short-term rental then it would have been if rented long-term over the course of the year.  This proportion, however, varied by location and property type.  While illustrative, the above do not fully control for property characteristics or owner actions -- which are a key component of short-term rental success.  To better understand which factors are related to a higher likelihood of short-term preference, we develop a set of logistic regression models below. -->

<!-- We specify three, increasingly complex, logistic regression models.  Model 1 takes the form of: -->

<!-- \begin{equation} -->
<!-- ln(\frac{STPref}{1-STPref}) = \beta_{0} + \beta_{1} PType + \beta_{2} BedBath + \epsilon -->
<!-- \end{equation} -->

<!-- where $StPref$ is the probability of short-term preference, $\frac{STPref}{1-STPref}$ is the log-likelihood of short-term prefernce, $PType$ is the property type (apt or house), $BedBath$ is the bedroom and bathroom combination of the property and $\epsilon$ is an error term. -->

<!-- For Model 2 we add the five geographic submarkets, $SubMrkt$ to better control for geographic variation in the probabilities.  -->

<!-- \begin{equation} -->
<!-- ln(\frac{STPref}{1-STPref}) = \beta_{0} + \beta_{1} PType + \beta_{2} BedBath + \beta_{3} SubMrkt + \epsilon -->
<!-- \end{equation} -->

<!-- Finally, in Model 3 we add three variables describing policies set by the Airbnb hosts: -->

<!-- \begin{equation} -->
<!-- ln(\frac{STPref}{1-STPref}) = \beta_{0} + \beta_{1} PType + \beta_{2} BedBath + \beta_{3} SubMrkt + -->
<!-- \beta_3 GPR + \beta_{4} MinStay + \beta_5 CancelPolicy +  \epsilon -->
<!-- \end{equation} -->

<!-- where $GPR$ is the total number of guests allowed per bedroom^[Hosts set the total number of guests.  We have normalized this by bedroom to avoid co-linearity in our model with the bedroom variables.], $MinStay$ is the minimum number of nights that a stay may have and $CancelPolicy$ is the difficulty with which a booking may be cancelled.  We have grouped the cancellation policies into three groups, Flexible (24 hour notice), Moderate (5 day notice) and Strict (50\% refund, 7, 30 or 60 day notice required). -->

<!-- The results of all three models are shown in Table 7. For $PType$ apartments are the reference category,  for $BedBath$, 2Bed/2Bath is the reference category, for $Submrkt$ the Core is the reference category and for $CancelPolicy$ moderate is the reference category.  As the additional variables are added, the model fit diagnostics steadily increase, suggesting that the increased complexity of the models adds to the quality of the fit.   -->

<!-- In Model 1, we see that when controlling for structural characteristics, Houses have a significantly lower probability of being more profitable in the short-term than apartments.  Small 1Bed/1Bath properties have a lower probability of being preferable in the short term than the reference class (2Bed/2Bath), while larger 3Bed/2Bath and 4Bed/2Bath have much higher probability of short term preference. -->

<!-- Adding the submarket variables (Model 2) shows that all areas have a much lower probability of short-term prefernce when compared to the core (the reference category) when controlling for structural characteristics.  These findings mirror those found above in Table 9.  The coefficients for the structural characteristics change little from Model 1 to Model 2, speaking the robustness of those estimates.   -->

<!-- Finally, in Model 3 we add in host behavior or policies. Allowing a greater number of guests per bedroom results in a higher probability of short-term preference.  This variable may be operating a proxy for additional sleeping spaces which allow for a higher nightly rate, ceteris paribus, to be charged.  The greater the minimum stay, the lower the probability of short-term preference.  Longer minimum stay may equate to lower occupancy rates as one or two night travelers are exempted from the property.  Owners with longer minimum stays may be looking to minimize costs such as cleaning and administration.  Flexible cancellation policies result in a lower probability of short-term preference, while a Strict policy has a weak, but positive relationship with short-term preference.  Strict policies do help maintain revenues, but may also scare off some potential guests.  Strict cancellation may be a luxury for the best and highly desired properties on the market and, within our model, may be acting as a proxy for some unobserved measure of quality or demand.   -->

<!-- ```{r, results='asis', echo=FALSE, message=FALSE, comment=FALSE, fig.cap='Logistic Model Results of Short Term Preference'} -->

<!--  names(full.table) <- c('Variable', 'Model 1', "Model 2", "Model 3") -->
<!--  x.full <- xtable(full.table, -->
<!--                   caption='Logistic Modeling Results\n') -->
<!--  align(x.full) <- 'lllll' -->

<!--  print(x.full,  -->
<!--        include.rownames=FALSE,  -->
<!--        hline.after=c(-1,0,15,20,20),  -->
<!--        caption.placement='top') -->


<!-- ``` -->


<!-- \pagebreak -->

<!-- ## Discussion -->

<!-- <!-- Although it gathers considerable interest in both economic and public policy circles, analyses of Airbnb market and their comparison to more traditional long-term leasing strategies remains relatively unaddressed.  This work attempts to fill this void.  --> -->

<!-- The Airbnb market is Melbourne is rapidly expanding, as total listings on the portal have jumped from around 7,000 in 2014 to more than 20,000 in August 2016.  However, of this total, only 6,600 or so of the properties are entire homes or apartments that were listed as available for at least one day during the September 1, 2015 to August 31, 2016 time frame.  Of these, only 4,400 are common bed/bath configurations with non-outlying rates and were booked for at least one night over the period of study. Overall, the Airbnb market is very small relative to the long-term market. We caution future researchers and policy analysts to look deeply at initial estimates of Airbnb counts as many of the listings on the market are either room shares and/or not active.     -->

<!-- Comparing annual revenues from the two leasing strategies shows that revenues from more traditional long-term rentals are, on average, higher and much less variable than those from the Airbnb market (Figure 10).  However, there is long right tail to the distibution of short-term revenues suggesting that some owners are doing better, and occasionally much better, on Airbnb than they would on the long-term market. This univariate comparison of revenues does not tell the entire story, as the Airbnb market does not have the same spatial distribution as the long-term market (Figure 8).  Short-term leasing is much more common in the central areas of the city, with small clusters showing up in suburban beach locations and rural tourist destinations like the Yarra Valley wine district.  -->

<!-- To control for these differences, we set out to make direct comparison of short and long term revenues on both the short and long term market observations. The shared information that is available on both sets of properties is limited. Our initial work found estimating long term rental values and tenant search times for the short term properties to be feasible with these data constraints while the opposite is much more difficult. In other words, there appear to by many unobserved variables (presense of wi-fi, parking and laundry facilities) and/or none-property specific characteristics (such as minimum stays, pet policies, number of allowed guests, cancellation policies) that influence short term nightly rates and occupancy rates.  As a result, we were not able to accurately estimate likely short-term rates for the long-term observations.   -->

<!-- In looking only at the short-term observation with estimated long-term rents we found that approximately one in three Airbnb properties brought in more revenue with its short term strategy than it would have with a long-term approach. As expected, this probability is not equal across space, property type or property size (bed and bath count).  Apartments are much more likely than houses to be more profitable as a short term rental.  Core city properties are more likely than other properties to prefer a short-term strategy, as are larger properties versus smaller ones.  For instance, 45\% of apartments in the city core generated more revenue on Airbnb than they likely would have on the long term market.  Examining only apartments with at least three bedrooms and two baths that figure increases to 60\%. -->

<!-- The nightly rates and total occupancies drive the profitability of short-term leasing.  As the nightly rate vs. occupancy rate heat maps and SVM plots (Figures 16, 17, 18 and 19) show, even with very high nightly rates, owners would need to achieve an occupancy rate of at least the median in market (around 40\%) in order to achieve a better than 50\% probability of being more profitable in a short-term strategy than a long term one. A 40\% occupancy rate means bookings at least three days out of every week all year long. In Melbourne summers this may be an easy feat as tourist numbers swell, but doing so on a cold week in July likely proves more difficult.    -->

<!-- The logistic regression models than we estimate show that most of the finding from the univariate and bivariate analyses done earlier are validated (Table 7).  Large apartments in the central city are the best bet for generating greater revenues through short term leasing.  We also test for the impact of host behavior on the probability of short term preference.  We find that more flexibility around the number of guests and the minimum stay increase the probability of short-term preference, while the reverse is true of cancellation policies (more strict equals higher short term preference probability).  As a host, it appears to be best to do what you can to attract bookings and then make it difficult for those bookings to be cancelled.     -->

<!-- ### Limitations and Future Research -->

<!-- There are a number of limitations to this research.  There are few shared property characterstics between the two datasets, which make cross-estimating rents and rates difficult, especially from the short-term market.  Part of the difficulty in estimating potential short-term rates for long-term properties is that the short term market appears to be heavily influenced by non-property specific factors such as the leasing policies (minimum stay, number of guests, etc.) themselves.  Future research should focus both on building better models to explain (and ultimately predict) short term rates and occupancies as well as to understand the impact of policies such as strict cancellation or no pets on the short term market.  It is likely that this work is being done proprietarily by Airbnb.com itself and other complementary firms such as AirDNA.com, our data provider.   -->

<!-- Additionally, we have opted to make comparisons between short and long-term strategies by cross estimating likely rates and rents.  Alternatively, a matching procedure could be employed that would match like properties between the two markets and use comparisons between the two as evidence of potential short and long-term revenues.   -->

<!-- In this work, we have made the assumption that, over the course of a year, the general maintenance costs (excluding basic janitorial) are identical between short and long-term leasing options.  We have done so as we were unable to location any evidence to suggest that one was higher than the other.  This may be a limitation of our study and could prove to be an avenue of future research.   -->

<!-- Beyond airBNB.com other short-term platforms exist (e.g. www.homeaway.com). This project doesn't take into account that individual properties may be advertised and rented out through different platforms. These are additional revenue streams that are not take into account and might increase the occupancy and revenue for short-term rentals. -->

<!-- Finally, our analysis looks at a single year, beginning in September (spring) and ending in August in which the property owner is given a binary option of either short or long term leasing strategy.  In reality, the possibilities may not be so rigid.  For example,given the fact that Melbourne's large university student population (many of whom are international) require housing during Autumn, Winter and Spring, while tourist swell in the summer months, an enterprising owner maybe able to pair long term leasing during the off-season with short-term leasing during peak times to maximize revenues.  In general, the rise of Airbnb and related portals give properties owners more options, and therefore more opportunities, to maximize returns on their property.  Future research should expand our scenario to look at more than just a single binary option over a fixed period.  We feel there may be significant opportunities for new ventures in industry to exploit the changing market for housing.    -->

\pagebreak

## References


<!-- Heat map with other characteristics on axis

<!-- ### Submarkets -->

<!-- ### Flexible Geo Analysis -->

<!-- ### Logit Models (maybe?) -->

<!-- ### Reference to Shiny App -->

<!-- ## Future Research -->

<!-- - Does time of purchase (which month) matter? -->
