---
title: "Data Analysis and Modeling"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
editor_options: 
  chunk_output_type: console
---

### Introduction

This document describes all of the code used to the generate the output included in the 'To Airbnb?' study. This analysis uses the output from the **3_data_prep.rmd** file found in the same online repository. 

## Preliminary Commands

&nbsp;

We begin with a number of preliminary commands.  First we load all of the necessary R libraries for this analysis. 

```{r load_libraries, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

  library(tidyverse)
  library(sf)
  library(geosphere)
  library(ggmap)
  library(multidplyr)
  library(tmap)

```

```{r set_options, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

  options(stringsAsFactors=FALSE)
  options(xtable.comment = FALSE)
  options(xtable.caption.placement='top')

```

Next, we source the necessary custom code files for the analysis.  

```{r source_files, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

  source(file.path(getwd(), 'functions', 'abb_Functions.R'))

```

### Set Parameters

```{r set_constants, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}
  
  # Exchange rate: AUS to US 
  exch_rate <- 1.32  

  # Date of investment decision
  decision_date <- lubridate::as_date('2015-09-01')
  
  str_col = col_df$rgb[1]
  
```

## Load Data

We then load the data.

```{r load_data, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE} 

  load(file.path(getwd(), 'data', 'prepared', 'prepared_data.RData'))

```

#### Compare Revenue Distributions

```{r flatten_ltr}

  ltr_df <- ltr_tdf %>% 
    unnest() %>%
    select(-suburb.x) %>%
    rename(suburb=suburb.y)

```

Before we examine the distributions of the observed short-term revenues and the estimated long-term revenues, we must produce the estimates of long-term revenue for the short-term properties.

We start by specifying the imputation model specification we will use to estimated likely long-term (weekly) rental prices for each short-term property.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  ltr_mod_spec <- formula(log(price) ~ as.factor(type) +
                            as.factor(bedbath) + as.factor(suburb) +
                            as.factor(month))

```

We then employ a custom function to impute the long term rents for the short-term properties.  The 'clip.field' is a spatial fixed effects that add spatial recognition to the models.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  str_tdf$tbg <- paste0(str_tdf$product,'_', str_tdf$geo_mrkt)
  ltr_df$tbg <- paste0(ltr_df$product,'_', ltr_df$geo_mrkt)

  imp_df <- abbImputeCompare(str.df=str_tdf, 
                                ltr.df=ltr_df, 
                                mod.spec=ltr_mod_spec, 
                                match.factor='suburb',
                                split.field='tbg', 
                                verbose=FALSE) 
  
  
```

From the imputed results, we extract the imputed rental values and merge to the original short-term data.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  # Extract out DFs (only those that imputed)
  str_tdf <- str_tdf %>%
    dplyr::left_join(imp_df, by='property_id') %>%
    dplyr::mutate(act_pref = ifelse(revenue_net > ltr_imp_revenue, 1, 0),
                  lik_pref = ifelse(lik_revenue_net > ltr_imp_revenue, 1, 0),
                  pot_pref = ifelse(pot_revenue_net > ltr_imp_revenue, 1, 0))

```


## Logit models

Now we attempt to explain the drivers of preference with a set of logistic regression models.  

First, a basic model with just structural characteristics.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  glm_lik_str <- glm(lik_pref ~ type + as.factor(bedbath),
                     family=binomial(link='logit'),
                     data=str_tdf)

  glm_pot_str <- glm(pot_pref~type+as.factor(bedbath),
                       family=binomial(link='logit'),
                       data=str_tdf)
```

Then a model that adds in the geographic submarkets

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  glm_lik_loc <- glm(lik_pref ~ type + as.factor(bedbath) + as.factor(geo_mrkt),
                     family=binomial(link='logit'),
                     data=str_tdf)

  glm_pot_loc <- glm(pot_pref ~ type + as.factor(bedbath)+ as.factor(geo_mrkt),
                     family=binomial(link='logit'),
                     data=str_tdf)
```

Finally a model that adds in host policies. 

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  glm_lik_host <- glm(lik_pref ~ type + as.factor(bedbath) + as.factor(geo_mrkt) +
                        I(max_guests/bedrooms) + min_stay + as.factor(canc_policy),
                      family=binomial(link='logit'),
                      data=str_tdf)

  glm_pot_host <- glm(pot_pref ~ type + as.factor(bedbath) + as.factor(geo_mrkt) +
                        I(max_guests/bedrooms) + min_stay + as.factor(canc_policy),
                      family=binomial(link='logit'),
                      data=str_tdf)
  
  lik_glms <- list(str=glm_lik_str,
                   loc=glm_lik_loc,
                   host=glm_lik_host)
  
  pot_glms <- list(str=glm_pot_str,
                   loc=glm_pot_loc,
                   host=glm_pot_host)
  
```

### Host types

Finally, we test to see if grouping by hosts types has any impact.  

First we must divide hosts up.  We start by looking at block rate.  
    
```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  ggplot(str_tdf, aes(x=block_rate)) + 
    geom_density(fill=str_col, color=str_col) +
    ylab('Frequency\n') + 
    xlab('\n% of Time Property is Blocked') +
    scale_x_continuous(breaks=seq(0,1,.2),
                       labels=c('0%', '20%', '40%', '60%', '80%', '100%')) +
      ggtitle('Host Block Rate Frequency\n') +
      theme(plot.title = element_text(hjust = 0.5)) +
      theme(axis.text.y=element_blank(),
          axis.ticks.y=element_blank())

```

We then add the number of blocked periods

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  ggplot(str_tdf, aes(x=block_rate, y=nbr_block * (366/total_days))) + 
    geom_point(fill=str_col, color=str_col, size=.2) +
    ylab('Nbr of Blocked Periods\n') + 
    xlab('\n% of Time Property is Blocked') +
    scale_x_continuous(breaks=seq(0, 1, .2),
                       labels=c('0%', '20%', '40%', '60%', '80%', '100%')) +
    ggtitle('Block Rate vs. Nbr. of Blocked Periods\n') +
    theme(plot.title = element_text(hjust = 0.5)) 

```

No real pattern emerges, so we divide manually. 

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  str_tdf$blockpertime <- round(str_tdf$nbr_block * (366 / str_tdf$total_days), 1)

  str_tdf$host_type <- 'Unknown'
  str_tdf$host_type[str_tdf$blockpertime > 12] <- 'Multi-Platform User'
  str_tdf$host_type[str_tdf$block_rate <= .25] <- 'Profit Seeker'
  str_tdf$host_type[str_tdf$block_rate >= .75] <- 'Opportunistic Sharer'
  str_tdf$host_type <- factor(str_tdf$host_type,
                              levels=c('Profit Seeker', 'Opportunistic Sharer', 
                                        'Multi-Platform User', 'Unknown'))

 ggplot(str_tdf, 
       aes(x=block_rate, y=blockpertime, color=host_type)) + 
  geom_point(alpha=.5, size=.8, shape=16) +
  ylab('Number of Blocked Periods\n') + 
  xlab('\n% of Time Property is Blocked') +
      ggtitle('Block Rate vs. Nbr. of Blocked Periods\n') +
 
    scale_x_continuous(breaks=seq(0, 1, .2),
                     labels=c('0%', '20%', '40%', '60%', '80%', '100%')) +
    scale_color_manual(values=c(col_df$rgb[2], col_df$rgb[4], col_df$rgb[7], 
                                col_df$rgb[9])) + 
    theme(legend.position='bottom',
           legend.title = element_blank(),
         plot.title = element_text(hjust = 0.5))+
     guides(colour = guide_legend(override.aes = list(size=3,
                                                     alpha=1)))

```

And re-run the logistic regression models. 

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  ht_. <- split(str_tdf, str_tdf$host_type)
  
  lik_spec <- as.formula(lik_pref ~ type + as.factor(bedbath) + as.factor(geo_mrkt) +
                          I(max_guests/bedrooms) + min_stay + as.factor(canc_policy))

  pot_spec <- as.formula(pot_pref ~ type + as.factor(bedbath) + as.factor(geo_mrkt) +
                          I(max_guests/bedrooms) + min_stay + as.factor(canc_policy))

  lik_ht_glms <- purrr::map(ht_.,
                         .f=glm,
                         formula=lik_spec,
                         family=binomial(link='logit'))
  
  pot_ht_glms <- purrr::map(ht_.,
                         .f=glm,
                         formula=pot_spec,
                         family=binomial(link='logit'))
    
```


```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  dir.create(file.path(getwd(), 'data', 'analyzed'))
  save(str_tdf, ltr_tdf, ltr_df, col_df, suburb_sf, studyarea_sf, sa1_sf, str_sf, ltr_sf,
       pot_glms, pot_ht_glms, lik_glms, lik_ht_glms, core_suburbs,
       file=file.path(getwd(), 'data', 'analyzed', 'analyzed_data.RData'))

```
