---
title: "Data Analysis and Modeling"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
editor_options: 
  chunk_output_type: console
---

### Introduction

This document describes all of the code used to the generate the output included in the 'To Airbnb?' study. This analysis uses the output from the **3_data_prep.rmd** file found in the same online repository. 

## Preliminary Commands

&nbsp;

We begin with a number of preliminary commands.  First we load all of the necessary R libraries for this analysis. 

```{r load_libraries, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

  library(tidyverse)
  library(sf)
  library(geosphere)
  library(ggmap)
  library(multidplyr)
  library(tmap)

```

```{r set_options, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

  options(stringsAsFactors=FALSE)
  options(xtable.comment = FALSE)
  options(xtable.caption.placement='top')

```

Next, we source the necessary custom code files for the analysis.  

```{r source_files, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

  source(file.path(getwd(), 'functions', 'abb_Functions.R'))

```

### Set Parameters

```{r set_constants, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}
  
  # Exchange rate: AUS to US 
  exch_rate <- 1.32  

  # Date of investment decision
  decision_date <- lubridate::as_date('2015-09-01')
  
```

## Load Data

We then load the data.

```{r load_data, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE} 

  load(file.path(getwd(), 'data', 'prepared', 'prepared_data.RData'))

```

### Calculating Net Revenues

We begin the analysis by calculating the net revenues for the short term data.  We calculate three types of net revenues, Actual, Likely (Extrapolated) and Potential. The differences are described more fully in the writeup.

Before we start, we estimate two figures necessary to determine utility costs: 1) Costs per person per day (cpppd) and the average number of guests per bedroom. The State of Victoria gives energy costs at \$330 (apartment) to \$520 (house).  For 2 persons in an apartment that works out to \$5 per person per day.  For a house of 4 persons that works out to about \$4 per person per day. We use a mid point value of \$4.50 per person per day.  We use an estimate of 1.5 persons per bedroom in each unit.

```{r utility_estimates, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  # Set cost per person per day
  cpppd <- 4.5

  # Calculate avg number of guests
  str_tdf$avg_guests <- str_tdf$bedrooms * 1.5

```

We then calculate the actual net revenue by multiplying booking times median rate - utility costs.  

```{r act_revenues, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  str_tdf$act_revenue <- ((str_tdf$bookings * str_tdf$med_rate) -
                                 (str_tdf$bookings * str_tdf$avg_guests * cpppd))


```

Next we create Likely Revenue, by imputing occupancy rates and revenues for the remainder of the year for those properties

```{r lik_revenues, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  # Create daily rate summary
  rate_summ <- str_tdf %>% 
   dplyr::select(property_id, daily_data) %>%
    tidyr::unnest() %>%
    dplyr::filter(status != 'B') %>%
    dplyr::group_by(date) %>%
    summarize(rate = length(which(status == 'R')) / n(),
            obs = n())

  # Create new cluster
  abb_cl <- get_default_cluster() 

  # Register function and variable
  cluster_copy(abb_cl, impLikelyBookings)
  cluster_copy(abb_cl, rate_summ)
  cluster_library(abb_cl, c("magrittr", 'dplyr', 'tidyr'))

  # Summarize daily by property
  lik_book_df <- str_tdf %>%
    dplyr::filter(total_days != 366) %>%
    multidplyr::partition(property_id, cluster=abb_cl) %>% 
    dplyr::do(impLikelyBookings(., std=g)) %>% 
    dplyr::collect()

  # Add to the existing data
  str_tdf <- str_tdf %>%
    left_join(lik_book_df, by='property_id') %>%
    dplyr::mutate(lik_bookings = if_else(is.na(lik_bookings), bookings, as.integer(lik_bookings)))

  # Calculate likely revenue
  str_tdf$lik_revenue <- ((str_tdf$lik_bookings * str_tdf$med_rate) -
                           (str_tdf$lik_bookings * str_tdf$avg_guests * cpppd))

  # Create Extrapolation parameter
  str_tdf$extr_par <- 366 / str_tdf$total_days

```

```{r pot_revenues, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  str_tdf$occ_rate <- 1 - (str_tdf$block_rate + str_tdf$avail_rate)

  str_tdf$pot_occ_rate <- ((str_tdf$lik_bookings) / (366 - (366 * str_tdf$block_rate)))

  str_tdf$pot_revenue <- ((str_tdf$pot_occ_rate * 366 * str_tdf$med_rate) -
                           (str_tdf$pot_occ_rate * 366 * str_tdf$avg_guests * cpppd))

```

#### Compare Revenue Distributions

```{r flatten_ltr}

  ltr_df <- ltr_tdf %>% 
    unnest() %>%
    select(-suburb.x) %>%
    rename(suburb=suburb.y)

```

Before we examine the distributions of the observed short-term revenues and the estimated long-term revenues, we must produce the estimates of long-term revenue for the short-term properties.

We start by specifying the imputation model specification we will use to estimated likely long-term (weekly) rental prices for each short-term property.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  ltr_mod_spec <- formula(log(price) ~ as.factor(type) +
                            as.factor(bedbath) + as.factor(suburb) +
                            as.factor(month))

```

We then employ a custom function to impute the long term rents for the short-term properties.  The 'clip.field' is a spatial fixed effects that add spatial recognition to the models.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  str_tdf$tbg <- paste0(str_tdf$product,'_', str_tdf$geo_mrkt)
  ltr_df$tbg <- paste0(ltr_df$product,'_', ltr_df$geo_mrkt)

  imp_df <- abbImputeCompare(str.df=str_tdf, 
                                ltr.df=ltr_df, 
                                mod.spec=ltr_mod_spec, 
                                match.factor='suburb',
                                split.field='tbg', 
                                verbose=FALSE) 
  
  
```

From the imputed results, we extract the imputed rental values and merge to the original short-term data.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  # Extract out DFs (only those that imputed)
  str_tdf <- str_tdf %>%
    dplyr::left_join(imp_df, by='property_id') %>%
    dplyr::mutate(act_pref = ifelse(act_revenue > ltr_imp_revenue, 1, 0),
                  lik_pref = ifelse(lik_revenue > ltr_imp_revenue, 1, 0),
                  pot_pref = ifelse(pot_revenue > ltr_imp_revenue, 1, 0))

```


## Logit models

Now we attempt to explain the drivers of preference with a set of logistic regression models.  

First, a basic model with just structural characteristics.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  glm_lik_str <- glm(lik_pref ~ type + as.factor(bedbath),
                     family=binomial(link='logit'),
                     data=str_tdf)

  glm_pot_str <- glm(pot_pref~type+as.factor(bedbath),
                       family=binomial(link='logit'),
                       data=str_tdf)
```

Then a model that adds in the geographic submarkets

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  glm_lik_loc <- glm(lik_pref ~ type + as.factor(bedbath) + as.factor(geo_mrkt),
                     family=binomial(link='logit'),
                     data=str_tdf)

  glm_pot_loc <- glm(pot_pref ~ type + as.factor(bedbath)+ as.factor(geo_mrkt),
                     family=binomial(link='logit'),
                     data=str_tdf)
```

Finally a model that adds in host policies. 

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  glm_lik_host <- glm(lik_pref ~ type + as.factor(bedbath) + as.factor(geo_mrkt) +
                        I(max_guests/bedrooms) + min_stay + as.factor(canc_policy),
                      family=binomial(link='logit'),
                      data=str_tdf)

  glm_pot_host <- glm(pot_pref ~ type + as.factor(bedbath) + as.factor(geo_mrkt) +
                        I(max_guests/bedrooms) + min_stay + as.factor(canc_policy),
                      family=binomial(link='logit'),
                      data=str_tdf)
  
  lik_glms <- list(str=glm_lik_str,
                   loc=glm_lik_loc,
                   host=glm_lik_host)
  
  pot_glms <- list(str=glm_pot_str,
                   loc=glm_pot_loc,
                   host=glm_pot_host)
  
```

### Host types

Finally, we test to see if grouping by hosts types has any impact.  

First we must divide hosts up.  We start by looking at block rate.  
    
```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  ggplot(str_tdf, aes(x=block_rate)) + 
    geom_density(fill=str_col, color=str_col) +
    ylab('Frequency\n') + 
    xlab('\n% of Time Property is Blocked') +
    scale_x_continuous(breaks=seq(0,1,.2),
                       labels=c('0%', '20%', '40%', '60%', '80%', '100%')) +
      ggtitle('Host Block Rate Frequency\n') +
      theme(plot.title = element_text(hjust = 0.5)) +
      theme(axis.text.y=element_blank(),
          axis.ticks.y=element_blank())

```

We then add the number of blocked periods

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  ggplot(str_tdf, aes(x=block_rate, y=nbr_block * extr_par)) + 
    geom_point(fill=str_col, color=str_col, size=.2) +
    ylab('Nbr of Blocked Periods\n') + 
    xlab('\n% of Time Property is Blocked') +
    scale_x_continuous(breaks=seq(0, 1, .2),
                       labels=c('0%', '20%', '40%', '60%', '80%', '100%')) +
    ggtitle('Block Rate vs. Nbr. of Blocked Periods\n') +
    theme(plot.title = element_text(hjust = 0.5)) 

```

No real pattern emerges, so we divide manually. 

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  str_tdf$blockpertime <- round(str_tdf$nbr_block * str_tdf$extr_par, 1)

  str_tdf$host_type <- 'Unknown'
  str_tdf$host_type[str_tdf$blockpertime > 12] <- 'Multi-Platform User'
  str_tdf$host_type[str_tdf$block_rate <= .25] <- 'Profit Seeker'
  str_tdf$host_type[str_tdf$block_rate >= .75] <- 'Opportunistic Sharer'
  str_tdf$host_type <- factor(str_tdf$host_type,
                              levels=c('Profit Seeker', 'Opportunistic Sharer', 
                                        'Multi-Platform User', 'Unknown'))

 ggplot(str_tdf, 
       aes(x=block_rate, y=blockpertime, color=host_type)) + 
  geom_point(alpha=.5, size=.8, shape=16) +
  ylab('Number of Blocked Periods\n') + 
  xlab('\n% of Time Property is Blocked') +
      ggtitle('Block Rate vs. Nbr. of Blocked Periods\n') +
 
    scale_x_continuous(breaks=seq(0, 1, .2),
                     labels=c('0%', '20%', '40%', '60%', '80%', '100%')) +
    scale_color_manual(values=c(col_df$rgb[2], col_df$rgb[4], col_df$rgb[7], 
                                col_df$rgb[9])) + 
    theme(legend.position='bottom',
           legend.title = element_blank(),
         plot.title = element_text(hjust = 0.5))+
     guides(colour = guide_legend(override.aes = list(size=3,
                                                     alpha=1)))

```

And re-run the logistic regression models. 

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  ht_. <- split(str_tdf, str_tdf$host_type)
  
  lik_spec <- as.formula(lik_pref ~ type + as.factor(bedbath) + as.factor(geo_mrkt) +
                          I(max_guests/bedrooms) + min_stay + as.factor(canc_policy))

  pot_spec <- as.formula(pot_pref ~ type + as.factor(bedbath) + as.factor(geo_mrkt) +
                          I(max_guests/bedrooms) + min_stay + as.factor(canc_policy))

  lik_ht_glms <- purrr::map(ht_.,
                         .f=glm,
                         formula=lik_spec,
                         family=binomial(link='logit'))
  
  pot_ht_glms <- purrr::map(ht_.,
                         .f=glm,
                         formula=pot_spec,
                         family=binomial(link='logit'))
    
```


```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  save(str_tdf, ltr_tdf, ltr_df, col_df, suburb_sf, studyarea_sf, sa1_sf, str_sf, ltr_sf,
       pot_glms, pot_ht_glms, lik_glms, lik_ht_glms, core_suburbs,
       file=file.path(data_path, 'analyzed', 'analyzed_data.RData'))

```
